Query Optimization for Dynamic Imputation

### 2017-05-15 paper/experiments.tex
some moves and highlight new text

The queries consist not only of projections and selections, but also
interesting joins and aggregates. Our aim was to craft meaningful queries that would
provide performance figures relevant to practitioners using similar datasets.

@@

We also calculate the number of tuples used to produce each aggregate output.
The count fraction columns in \Cref{table:smape} show the number of tuples in the aggregate for $\alpha = 0$ and $\alpha = 1$ as a fraction of the number of tuples used when running the query on the imputed base table.
This shows that when optimizing for performance, not quality, many tuples are dropped.
Even in cases where the SMAPE reduction between $\alpha = 1$ and $\alpha = 0$ is small (\Cref{q2}, \Cref{q6}) the tuple count is significantly different.
In these cases, the aggregate value is not significantly impacted by the missing data. In particular, if values are missing uniformly at random,
the aggregate should not be affected.
However, if the missing data is biased then the aggregate will have a significant error.
This highlights a challenge for a user handling data imputation traditionally: it is unclear if the missing data will have a large or small negative impact on their analysis until they have paid the cost of running it.
By using \ProjectName{} this cost can be lowered significantly. We can also trivially extend \ProjectName{} to warn users when the query chosen for execution has a high
$\textsc{Penalty}$ estimate, along with the number of tuples that have been dropped (after execution), so that situations with high potential for skewed results can be identified by the user.

Our benchmark queries performed well for both cases of \texttt{AVG} and \texttt{COUNT} aggregates, so we expect \ProjectName{} would provide similar performance for \texttt{SUM}, with 
errors proportional to the fraction of relevant tuples retrieved and the \texttt{AVG} value estimated.

\review{
\Cref{fig:pareto-frontiers} shows the final Pareto frontiers produced by the planner for \Cref{q1} and \Cref{q6}. These are the resulting set of plans after pruning dominated plans throughout our algorithm.
and encapsulate the possible trade-offs that the user can make in choosing a final plan for execution. The frontier is available at the end of the planning stage and
can be exposed to the user to help in guiding their workflow.
}

@@

In contrast, \ProjectName{} executes the query \lstinline{SELECT AVG(c0) FROM acs} over the ACS dataset (also using chained-equation decision trees imputation) in \acsimputedbzeroresult{} when optimizing for quality and \acsimputedboneresult{} when optimizing for runtime. Given the runtime difference, we can run approximately 5,000 queries before taking more time than full imputation.
The performance difference comes from the fact that \ProjectName{} only needs to perform imputation on the single column required by the query.
Adding selection predicates would further reduce the query runtime by reducing the amount of required imputation.
An analyst could do the same, but tracking the required imputations would get increasingly complicated as the queries became more complex.
This highlights the benefit of using our system for early data exploration.

>>>>

The queries consist not only of projections and selections, but also
interesting joins and aggregates. Our aim was to craft meaningful queries that would
provide performance figures relevant to practitioners using similar datasets.
\review{
Our benchmark queries performed well for both cases of \texttt{AVG} and \texttt{COUNT} aggregates, 
so we expect \ProjectName{} would provide similar performance for \texttt{SUM}, with 
errors proportional to the fraction of relevant tuples retrieved and the \texttt{AVG} value estimated.}

@@

We also calculate the number of tuples used to produce each aggregate output.
The count fraction columns in \Cref{table:smape} show the number of tuples in the aggregate for $\alpha = 0$ and $\alpha = 1$ as a fraction of the number of tuples used when running the query on the imputed base table.
This shows that when optimizing for performance, not quality, many tuples are dropped.
Even in cases where the SMAPE reduction between $\alpha = 1$ and $\alpha = 0$ is small (\Cref{q2}, \Cref{q6}) the tuple count is significantly different.
In these cases, the aggregate value is not significantly impacted by the missing data. In particular, if values are missing uniformly at random,
the aggregate should not be affected.
However, if the missing data is biased then the aggregate will have a significant error.
This highlights a challenge for a user handling data imputation traditionally: it is unclear if the missing data will have a large or small negative impact on their analysis until they have paid the cost of running it.
\review{
By using \ProjectName{} this cost can be lowered significantly. We can also trivially extend \ProjectName{} to warn users when the query chosen for execution has a high
$\textsc{Penalty}$ estimate, along with the number of tuples that have been dropped (after execution), so that situations with high potential for skewed results can be identified by the user.

\Cref{fig:pareto-frontiers} shows the final Pareto frontiers produced by the planner for \Cref{q1} and \Cref{q6}. These are the resulting set of plans after pruning dominated plans throughout our algorithm.
and encapsulate the possible trade-offs that the user can make in choosing a final plan for execution. The frontier is available at the end of the planning stage and
can be exposed to the user to help in guiding their workflow.
}

@@

In contrast, \ProjectName{} executes the query \lstinline{SELECT AVG(c0) FROM acs} over the ACS dataset (also using chained-equation decision trees imputation) in \acsimputedbzeroresult{} when optimizing for quality and \acsimputedboneresult{} when optimizing for runtime. 
\review{Given the runtime difference, we can run approximately 5,000 queries before taking more time than full imputation.
The performance difference comes from the fact that \ProjectName{} only needs to perform imputation on the single column required by the query.
Adding selection predicates would further reduce the query runtime by reducing the amount of required imputation.
An analyst could do the same, but tracking the required imputations would get increasingly complicated as the queries became more complex.
This highlights the benefit of using our system for early data exploration.}
